{\bf\fontsize{14pt}{18pt}\selectfont AICov: An Integrative Deep Learning Framework for COVID-19 Forecasting with Population Covariates}

### 2020-12-18 response-1.tex

\documentclass{article}
\usepackage{fullpage}
\usepackage[utf8]{inputenc}

\title{Review Response JDS2007-001RA0}

\date{Revision due: 2020-11-24}

\begin{document}

\maketitle

\parindent 0pt We are happy to resubmit our paper and are very thankful for the great reviews we received. Thanks to the review comments, the quality of the paper has significantly improved.

In particular, we addressed the following:

\begin{enumerate}

\item The paper has been checked multiple times by multiple native English speakers.

\item The organization of the paper has been improved while focusing entirely on the deep learning framework. Several sections and images that distracted the reader from the paper's primary focus (which was evident by the careful reviews) were removed.

\item Reviewer 1: The method has been described using a recipe approach without giving the rationality behind it. There are too many predictive models for COVID-19. What are the advantages of the proposed method? The authors should explain their competitiveness, at least from the following two perspectives. We addressed these issues through the next two items.

\item Reviewer 1: First, compared with the very classical SIR/SEIR and their improved models, what are the advantages of the proposed model?

    A new section has been added to discuss the advantages and disadvantages.

\item Reviewer 1: Is the accuracy improved significantly?

    While using the real-time data our accuracy is very good as it, for example accounts for the weekly fluctuations that are not covered by other prediction models. Moreover, the point is not only a single model in time, but that the model adapts itself over time while new data is added. This has been pointed out in the paper.

\item Reviewer 1: Second, it seems that the data used are all very structured, what about regression model framework?

    The beauty of this deep learning algorithm is that it learns the trends and performs well even if the data is not smoothed with regression or a seven-day average. Instead, it even learns the weekly fluctuations which we point out in the revised version of the paper.

\item Reviewer 1: Is the performance competitive?

    The performance is very competitive and an explanation is added to the conclusion.

\item Reviewer 1: The current deep learning framework is lack of interpretation of the investigated risk factors.

    We added the following section:

While this study focuses on if and which risk factors can improve the predictions, we do not focus on why. Such an analysis could provide insights into causal relationships. For example,  information such as health insurance and the number of hospitals and beds are often related to the quality of services that affect early detection and treatment. Furthermore, population parameters such as density can determine the rate of spread and the number of cases. Similarly, distributions of comorbidities in a given population such as physical health, cholesterol screening, diabetes, etc., are found to be insightful in predicting outcomes. Identifying such risk factors and correlating them to the prediction accuracy provides valuable candidates for future studies.

\item Reviewer 2: The manuscript is in a rather preliminary shape, and needs significant work before being in a publishable shape.

    We believe we have addressed this issue by addressing all reviewer's comments and more.

\item Reviewer 2: Section 2.2 does not look sufficient as a subsection.  It should be either expanded to include more details, or combined with Section 2.1.3.

    The sections have been improved/combined, and a section from later on in the paper was added.

\item Page 5, under "Data Source Integration", what does "we include integrate data integration capabilities" mean?

    The sentence has been corrected.

\item Reviewer 2: 4.  Please consider writing the requirements on Pages 5 and 6 into structured paragraphs.

    The requirements have been rewritten as paragraphs. Also, we added a numbered requirement tag to them and kept the high-level name of the requirement for increased readability and structure.

\item Reviewer 2: Item 5., 6., 7., 8., 9., and 10. Have been fixed by excluding the text related to them from the paper as we found they distracted too much from our main contributions.

\item Reviewer 2: 12.  In Section 2.1,  the activation function is the sigmoid function.  In Section 6, RELU was used.

    Thanks for detecting it. We used in both approaches sigmoid only in the last layer and corrected it in the paper

\item Reviewer 2: 13.  Please consider adding labels for the yaxes in Figure 10 and Figure 11.

    All images now have information about the axes.

\item Reviewer 2: 14.  Page 19, please fix the broken link to reference.

    Broken links were fixed.

\item Reviewer 2: 15.  Sections 7 and 8 are rather redundant.  The language does not read clear.  I suggest moving these two sections, and make Section 5 and 6 a more complete story.

    Thanks for this suggestion. To better showcase our arguments we have done the following: (a) most of the theory was moved into the same section in order to reduce redundancy, (b) we introduced a number of Questions that we analyze within this report to provide additional structure and delineate the sections which answer each question, (c) a more comprehensive explanation has been added the later sections. Together they provide a more complete story.

\item Reviewer 2: 16.  Please consider making the code publicly available at GitHub.

    We intend to make the code public after publication. More than 30K lines of underlying infrastructure code and documentation is already available on GitHub and citations have been added.

\end{enumerate}

\end{document}

### 2020-12-18 jds-paper-v3.tex
add overleaf paper to github

\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{array}
\usepackage{authblk}
\usepackage{color}
\usepackage[table]{xcolor}
\usepackage{epsfig,fancyhdr}
\usepackage{geometry,graphicx}
\usepackage[]{hyperref} 
\usepackage{indentfirst}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{listings}
\usepackage{natbib}
\usepackage{rotating}
\usepackage{comment}
\usepackage{booktabs}
\usepackage{todonotes}
\setuptodonotes{inline}

@@

\setcounter{page}{1}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{remark}{Remark}
\pagestyle{empty}

\setcitestyle{square}

@@

%\pagestyle{fancy}
%fancyhead[RO]{odd number page}
%fancyhead[LE]{even number page}

@@

@@

@@

\affil[1]{{\emph\fontsize{12pt}{0.5em}\selectfont Digital Science Center, Indiana University, Bloomington Indiana, USA}}
\affil[2]{{\emph\fontsize{12pt}{0.5em}\selectfont Public Health Dynamics Lab, and Department of Biostatistics, University of Pittsburgh, Pittsburgh, Pennsylvania, USA}}
\affil[3]{{\emph\fontsize{12pt}{0.5em}\selectfont Health Analytics Network, Pennsylvania, USA}}

\newcommand{\ABSTRACT}{
The COVID-19 pandemic has had profound global consequences on health, economic, social, behavioral, and almost every major aspect of human life. Therefore, it is of great importance to model COVID-19 and other pandemics in terms of the broader social contexts in which they take place. 
We present the architecture of AICov, which provides an integrative deep learning framework for COVID-19 forecasting with population covariates, some of which may serve as putative risk factors. 
We have integrated multiple different strategies into AICov, including the ability to use deep learning strategies based on LSTM and event modeling.
To demonstrate our approach, we have introduced a framework that integrates population covariates from multiple sources. Thus, AICov not only includes data on COVID-19 cases and deaths but, more importantly, the population's socioeconomic, health, and behavioral risk factors at a local level. The compiled data are fed into AICov, and thus we obtain improved prediction by the integration of the data to our model as compared to one that only uses case and death data. As we use deep learning our models adapt over time while learning the model from past data.
}
\date{}

@@

During outbreaks of epidemics, initially data on cases and deaths could be scarce, and the quality of data annotation, validation, and aggregation might be uncertain. For instance, changes in clinical data entry, such as the addition of a new category {\em clinically diagnosed} to the existing {\em lab-confirmed} category, may likely have been reflected in the reporting \cite{Petropoulos2020-hh}. Initial fears about COVID-19 among the general population with regards to the trajectory of the pandemic could also affect administrative reporting. Diverse media sources and differences in local and federal policies may add to the general uncertainty about disease progression. Nevertheless, a data-driven approach to forecasting can offer valuable insights into the disease dynamics, and thereby an ability to objectively plan for the near future.

Unlike earlier global outbreaks, during COVID-19, the current worldwide digital ecosystem allows for real-time data collection, which in turn is used by AI and deep learning systems to understand healthcare trends, model risk associations, and predict outcomes \cite{Ting2020-lw}. 
In addition to traditional public health surveillance strategies available to most countries, a variety of static and dynamic data types may be integrated to model the scale and dynamics of this pandemic \cite{Pyne2015-ao}. Several organizations, including the Johns Hopkins University’s Center for Systems Science and Engineering, the New York Times, the Atlantic’s COVID-19 tracking project, among others have developed real-time tracking maps for following COVID-19 cases around the world using data from the U.S.~Centers for Disease Control and Prevention (CDC), WHO, and other international health agencies. Notably, the CDC receives forecasts from several modeling groups that use a wide variety of approaches ranging from SEIR (Susceptible-Exposed-Infectious-Recovered) to Agent-based to Bayesian models to understand the non-pharmaceutical policy interventions (or the lack thereof) to predict disease dynamics and its impact on human lives \cite{www-cdc-modeling-forecast}.

For precise and contextually relevant modeling of COVID-19 dynamics for a given population or community, we use data integration to combine community-specific health and behavioral risk factors, demographic and socioeconomic variables that are available at the local levels along with the cases and deaths data for COVID-19 while leveraging the recently developed and increasingly popular deep learning strategies to forecast COVID-19 dynamics using our model. Towards this, we develop a parallel computing platform called AICov standing for  {\em AI-driven Platform for COVID-19} to implement the different components of our framework and obtain the needed resources through multi-cloud interfaces in a robust manner.

In this study, while we incorporate information from community and county-specific covariates to inform our forecasting model for each metropolitan area, we want to avoid the so-called ecological fallacy in drawing inferences about individual disease outcomes based on large area-level aggregated risk factors. Our objective is also to underscore the nuanced roles played by pre-existing socioeconomic and other prevailing conditions of a given community that can act as determinants of its health outcomes and possible disparities,  especially under such sudden stresses to its current systems as those felt during any pandemic.
It is important to note that the underlying models are {\em learned} directly from the data instead of being projected by a theoretical epidemiological model that relies on simulation populations. This allows quick adaptation and self-adaptation in regards to changing behaviors and regional differences.

@@

\begin{comment}
(Section \ref{sec:lstm-background}).
Additionally, we provide a short introduction of the LSTM implementation as used in Keras (Section \ref{sec:keras}) that we used as a foundation of this work.
\end{comment}

@@

\begin{comment}
% ------------------------------
\subsection{LSTM in Keras}
\label{sec:keras}
% ------------------------------

LSTM has been available in Keras since 2015 \cite{www-keras-lstm,www-keras-api}. However, it does require some level of familiarity with RNNs for the required manipulation of the input data which needs to be adjusted for the application. The lack of complex and general examples involving multiple input and output variables, in conjunction with secondary information, calls for special attention from the users. Most examples focus on very simple usage of LSTM and do not to take into account the integration of other variables, such as variables represented by spacial population risk factors, in addition to time series data as input and output used in our approach. 
A further issue is that Keras depends on and is distributed with Tensorflow \cite{www-tensorflow}. Despite the popularity of Tensorflow as a Deep Learning framework, its adoption by modern operating systems has lagged behind other frameworks such as PyTorch \cite{www-pytorch}. 
Our framework overcomes these important limitations, especially while addressing the integration of spacial variables that could pose additional input for improving the accuracy of the predictions.
\end{comment}

@@

\begin{requirement}{\bf Expandable API.} Assuming that new models and other analysis algorithms will be developed over time, AICov must allow the integration of such APIs. \Solution To address this requirement, we are developed an abstraction API for data, analysis, and metadata parameter adjustments. External services could be integrated with the help of REST services. The REST services are automatically generated from function specifications. 
\end{requirement}

@@

Among the different sources of data on daily cases and deaths data collected and aggregated at different spatial levels, especially for the US, many are either freely available repositories such as GitHub or accessible via specifically developed online resources such as the New York Times~\cite{github-nytimes} and the Johns Hopkins Coronavirus Resource Center~\cite{www-jh-covid19}. 
For this study, we generated two data sets based on the latter resource upon our checking for the locations and times at which cumulative counts of COVID-19 cases and deaths on any given day were no less than those on the previous day. This led us to compile daily coronavirus cases and deaths data for 345 US counties from 32 states and the District of Columbia and matched by five-digit FIPS code or county name to dynamic and static variables from additional data sources.

The static data in this repository was collected from the US Census Bureau and the Centers for Disease Control and Prevention (CDC). The different fields cover the domains of behavioral and health risk factors, hospital capacity, and socioeconomic and demographic conditions. Data on health outcomes, prevention measures, and unhealthy behaviors were extracted from the CDC's 500 Cities Local Data for Better Health program \cite{www-cdc-chronic-data} in the form of crude or age-adjusted prevalence of conditions such as respiratory disease, obesity or smoking. These values were aggregated from the census tract to the county level using the first five digits of the eleven-digit tract FIPS code. Because a single county may consist of multiple individual cities, we include the list of all city labels within each aggregate group to represent a greater metropolitan area. 110 of such metropolitan areas that had more than 500 reported cases of COVID-19 by April 15, 2020, were selected for this study.

Notably, a greater metropolitan area such as New York City could be spread across more than one county, which can lead to the complexity of aggregating counts and other variables. Time series data for the 5 individual counties (boroughs) in New York City were not available in the Johns Hopkins CSSE data set. Rather, the total number of cases and deaths for the entire city of New York are reported and assigned the FIPS code of New York County (Manhattan). To ensure consistent geography across the static and dynamic data, we compute the population-weighted sum or median of each covariate over the assigned counties.

@@

\begin{comment}
We are using the following abbreviations in the paper, and the data with data from \cite{cdc-brfss}, 
\cite{cdc-svi} and \cite{www-500Citie16} of population $>=18$ years of age that have been data fused by University of Pittsburgh.
\end{comment}

\begin{table}[!hptb]
\caption{Covariate Risk factor abbreviations as used in this study.}
\label{tab:risk-factors}
\bigskip
\resizebox{0.90\textwidth}{!}{
\centering
\begin{tabular}{ll}
\toprule
Risk Factor & Description \\
\midrule
 \underline{None} & no risk factor used\\
 ARTHRITIS 	  & Percent of arthritis \\
 BINGE 		  & Percent of binge drinking \\
 BPHIGH 	  & Percent of high blood pressure\\
 BPMED 		  & Percent of taking blood pressure medication\\
 CANCER 	  & Percent of cancer patients\\
 CASTHMA 	  & Percent of current asthma\\
 CHD 		  & Coronary heart disease\\
 CHECKUP 	  & Percent of health checkup\\
 CHOLSCREEN   & Percent of cholesterol screening\\
 COPD 		  & Percent of Chronic obstructive pulmonary disease\\
 CSMOKING 	  & Current smoking in percent\\
 DIABETES 	  & Percent of diabetes\\
 Estbeds 	  & Number of estimated beds in hospitals\\
 HIGHCHOL 	  & Percent of high cholesterol\\
 Insurance 	  & Percent of Insurance\\
 KIDNEY 	  & Percent of kidney disease\\
 LPA 		  & Percent of No leisure-time physical activity \\
 MHLTH 		  & Percent of not good mental health\\
 Nbeds 		  & Number of beds in hospitals \\
 Nbeds\_per1000   & Number of beds per 1000 people\\
 Nhosp 		  & Number of hospitals\\
 OBESITY 	  & Percent of population that is obese\\
 PHLTH        & Percent of Physical Health\\
 PVI 		  & Pandemic vulnerability index \\
 Percentblacks 	  & Percent of Blacks in the population\\
 Percenthispanics & Percent of Hispanics in the population\\
 STROKE 	  & Percent of stroke\\
 black\_percent	  & Percent of blacks in the population\\
 poverty\_percent & Percent of poverty in the population\\
 senior\_percent  & Percent of seniors in the population\\
 svi\_minority 	  & Percent of social vulnerability index in the minority population\\
 svi\_overall 	  & Percent of social vulnerability index in the overall population\\
\bottomrule
\end{tabular}
}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% COMMENTED OUT
%
\begin{comment}

\subsection{Covariance Analysis}

For each FIPS we have a number of risk factors for which we have values. These values can be correlated with each other through a correlation analysis.

To provide an initial idea how the risk factors are correlated with each other we showcase in Figure \ref{fig:cov-heatmap-1}  and \ref{fig:cov-heatmap-2} in the scale of -1.0 to 1.0. While the first figure shows all variables, we display in the second figure, only the variables related to health measurements and not social-economic factors.

In addition, we show in Figure \ref{fig:cov-pairplot} the more detailed pair plots. However, an online version of this plot is available, as indicated in the caption, as the information for the later is very condensed.

We observe that most social-economic variables are more strongly correlated with each other than the health variables. This is evident from the plot due to its coloring that is darker than the rest of Figure \ref{fig:cov-heatmap-1}. Obviously, the number of available hospitals vs. the number of estimated beds and the population are strongly correlated.

When looking at the health-related risk factors, we, for example, see that diabetes and high blood pressure are correlated with each other. This reflects observations from health care providers that have correlated these two variables. A similar correlation we see in the correlation between obesity and diabetes. Having such detailed information at hand makes it now suitable to integrate such health-related factors to COVID-19 cases and death.  This is conducted through two different comprehensive and sophisticated analysis studies.

In the first, we relate the risk factors based on information we predict for the maximum of the cases and death from the model we used for the given time period. The second is when we use our model predictions and find the best predictions while minimizing the cumulative error for the given prediction time period. To be complete, we have included all risk factors in our analysis rather than to restrict them based on our correlation matrix.

>>>>

\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{array}
\usepackage{authblk}
\usepackage{color}
\usepackage[table]{xcolor}
\usepackage{epsfig,fancyhdr}
\usepackage{geometry,graphicx}
\usepackage[]{hyperref} 
\usepackage{indentfirst}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{listings}
\usepackage{natbib}
\usepackage{rotating}
\usepackage{comment}
\usepackage{booktabs}
\usepackage{todonotes}
\setuptodonotes{inline}
\usepackage{makecell}
\usepackage{enumitem}
\usepackage{tcolorbox}

@@

\setcounter{page}{1}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{remark}{Remark}
\pagestyle{empty}
\usepackage{lineno}
\setcitestyle{square}

\newcommand*\rot{\rotatebox{90}}

@@

%\fancyhead[RO]{odd number page}
%\fancyhead[LE]{even number page}

\pagestyle{plain} 
\linenumbers

@@

\newtcolorbox{redbox}{colback=red!5!white,colframe=red!75!black,width=3cm}
\newtcolorbox{yellowbox}{colback=yellow!5!white,colframe=yelloe!75!black,width=3cm}

\newcommand{\DONE}{{\color{green!60!black}\makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{0.1em}$\checkmark$}}~}
\newcommand{\DOIT}{{\color{red!60!black}\makebox[0pt][l]{$\square$}\raisebox{.15ex}{\hspace{0.1em}$\boxtimes}}~}
\newcommand{\PROGRESS}{(in progress)~}

@@

REVISION DUE: 01.12.2021

\section*{Reviewer 1}

My comments are the following:

The revised manuscript is much smoother than the first version, but I still feel that less than sufficient care was put into the writing. The reviewer pointed some in the pdf report. I have some additional editorial suggestions:

\begin{enumerate}

\item \PROGRESS Submit the point-by-point response to the reviewer's comments and my comments in a pdf file.

\begin{quote}
The original submisison web page had a field to enter the comments as text which we used. We will change this and introduce a PDF file instead, which is easier for us also.
\end{quote}

\item \DONE Needs page number (line number) for ease of referencing in the review.

\begin{quote}
Page numbers and line numbers were activated
\end{quote}

\item \DONE Do not italicize in the text.

\begin{quote}
All italicized text have been removed.
inline formulas are done by latex italized. 
\end{quote}

\item \DONE AICov needs to be defined in the abstract.

\begin{quote}
We added

"We present the architecture of an artificial intelligence enhanced COVID-19 analysis (in short AICov)"

was added to the abstract.
\end{quote}

\item \PROGRESS Acronyms need to be defined at their first occurrences (e.g., AI in the intro). An acronyme table is included in table 1, we added the following acronyms to the paper

\begin{quote}
    All acronyms have been defined at their first usage. In addition we have added a section Acronyms in which all acronyms are listed to make it easier for the reader to look them up in case they are used later on in the paper.
\end{quote}

\DOIT capitalization in definition sin the table 1 and others

\DOIT PERCENT, PERCENTBALCK, BLACK PERCENT, also for hispanics and senior

\DOIT use risk factor in table headings

\DOIT PER1000 is /1000

\DOIT make acronyms appear in text

\DOIT Table 1. capitalization,explain why they are different or make them the same

\item \DOIT  Figure 2: please present the equations as equations. The notations used in the equation need a full check (e.g., h is used as both the dimension and vector). I'd like to see a clear distinction between inputs/outputs and parameters. I'd suggest that the inputs be set up before the equations are presented.

\begin{quote}
    \todo{todo}
\end{quote}

\item \DOIT  Regenerate the figures into pdf format (the current jpg/png figures blur when zoomed in).

\begin{quote}
    \todo{todo}
\end{quote}

\item \DONE  Table 2: consider reformating it; it would not fit the page width in print.

\begin{quote}
Table 2 has been reformatted. It was previously fit in the column with resize box. While rotating the headers the font is increased. The table still fits in the column as it still uses resizebox.
\end{quote}

\item \DONE  Table 3: tidy the headers.

\begin{quote}
    The headers have been cleaned up and linebreaks wer introduced so the resizebox of the table presents the table now larger
\end{quote}

\item \DOIT  Figure 5-6: can be combined to one figure with two panels (using ggplot2 and save the space of the labels of the horizontal axis).

\begin{quote}
    \todo{todo}
\end{quote}

\item \DOIT  Figure 7-8: can be one figure with two panels; pop density 2010 needs to be cleaned.

\begin{quote}
    \todo{todo}
\end{quote}

\item \DOIT  Figure 9-14: combine into one figure.

\begin{quote}
    \todo{todo}
\end{quote}

\item \DONE  Table 4: keep the number of digits consistent.

\begin{quote}
    The digits have been padded with the missing 0 to make them all the same length
\end{quote}

\item \DOIT  Reproducibility: submit data/code/README in a compressed file as supplementary material, or put them in a public repo and give the link under an unnumbered section Supplementary Material.

\begin{quote}
    \todo{todo}
\end{quote}

\end{enumerate}

To submit your revision, please log in to EJMS and submit it as a revised file to the original submission. Please also include a detailed description of how you addressed all the points raised by the reviewers.

\section*{Reviewer 2}

This paper presents AICov, an integrated deep learning framework for COVID-19 forecasting. It takes into account population covariates, and uses LSTM and event modeling to do the forecasting. Data integration is done at a local level. Integration of such information brings improvement to the prediction accuracy. Despite the authors’ effort to address my comments in their revision, the paper is still in rough shape. Please see my comments.

\begin{enumerate}

\item \DOIT Page 5, please write the full name for REST (representational state transfer ) when it first appears.

\begin{quote}
    The term has been defined at first usage, we also checked all other abbreviations.

    \todo{check other abbreviations}

\end{quote}

\item \DOIT Page 5, line 3 under Requirement 3, “we are developed an abstraction API for data” does not read right.

\begin{quote}
    \todo{todo}
\end{quote}

\item \DOIT Page 6, line 8, what level is “local level”?

\begin{quote}
    \todo{todo}
\end{quote}

\item (partially done) Page 7, Table 1, is “CHD” also a percent? If yes please modify the description; Percent- blacks and black percent both correspond to “Percent of blacks in the population”. Both seem to have been used according to Table 3, with difference in the four model performance measurements. Do the authors have an explanation for this?

    \begin{quote}
    The table has been updated and it was indicated that it is in percent. There were indeed to different information sources that reported different values for percentage of blacks and other populations. This however has since than corrected and we have been able to use just one population value.

    \todo{fix the values and only have one population value}
    \end{quote}

\item \DOIT Page 8, line 5 of Section 5.1, the abbreviation LSTM has already appeared before, and the full name can be removed.

    \begin{quote}
        \todo{todo}
    \end{quote}

\item \DOIT What do the two panels in Figure 4 represent? Please give clearer description either in the text or figure caption.

    \begin{quote}
        \todo{todo}
    \end{quote}

\item \DOIT Section 5.2, paragraph 2 line 6, no need to capitalize for the word “Samples”. Line 5 from the bottom, no need to capitalize for the word “Insurance”. Please be careful with your writing.

    \begin{quote}
        \todo{todo}
    \end{quote}

\item \DOIT Section 5.2, the “order for cases sorted by minimum model order” and “order for deaths sorted by minimum model error” paragraphs are redundant, as the same information is already displayed in the x-legend for Figures 5 and 6.

    \begin{quote}
        \todo{todo}
    \end{quote}

\item \DOIT In Table 3, what is the unit for the error terms? The errors in Figures 7 and 8 are percentages. If errors in Table 3 are also percentages please clearly mark. Also, the numbers in Figures 7 and 8 do not align with numbers in row 1 of Table 3. Please check your numbers with care.

    \begin{quote}
        \todo{todo}
    \end{quote}

\item \DOIT Please adjust the location of your figures so they do not look pages apart from their interpretations.

    \begin{quote}
        \todo{todo}
    \end{quote}

\item \DOIT If I am reading it correctly, Figures 9 and 11 are only the “zoom in” versions of partial information from Figure 13. Similarly for the right column. Why waste space on two more figures when you could make Figures 13 and 14 bigger and explain the findings?

    \begin{quote}
        \todo{todo}
    \end{quote}

\item \DOIT The last paragraph of Section 5.2 suits the conclusion/discussion for future work section, not in the middle of a paper.

    \begin{quote}
        \todo{todo}
    \end{quote}

\item \DOIT Section 5.3, the authors mentioned two way of constructing covariates. Since the second way was more complicated while yielding no better performance the authors did not use it. Then it can be removed without damaging the storyline.

    \begin{quote}
        \todo{todo}
    \end{quote}

\item \DOIT The description “We renormalized, so the two basic features (following day) contributed 50% of the loss function” is hard to interpret. Please elaborate.

    \begin{quote}
        \todo{todo}
    \end{quote}

\item \DOIT Paragraph 2 on page 14, “Selu, Relu and Tanh activations gave very similar results while Sigmoid activation was much worse. However we still use a Sigmoid function...” is very counter-intuitive. Isn’t the point of trying different activation functions to better fit the data?
16. Paragraph 3 on page 14, “The basic COVID-19 daily data is extensive but compared two approaches with and without renormalization, but in this case, we ...” The sentence does not read correct. Please check your sentences with care!

    \begin{quote}
        \todo{todo}
    \end{quote}

\item \DOIT How to interpret the “significant positive correlation” in Section 6 from the practical perspec- tive? Also, the term “evolution function” made its first appearance in Section 6 without any definition or explanation; I suggest either removing the whole Section 6, or extend it with clearer definition of evolution function.
18. The first two paragraphs do not read like to be in a “Conclusion” section. Rather, they better fit into the introduction.

    \begin{quote}
        \todo{todo}
    \end{quote}

\end{enumerate}

@@

\affil[1]{{\fontsize{12pt}{0.5em}\selectfont Digital Science Center, Indiana University, Bloomington Indiana, USA}}
\affil[2]{{\fontsize{12pt}{0.5em}\selectfont Public Health Dynamics Lab, and Department of Biostatistics, University of Pittsburgh, Pittsburgh, Pennsylvania, USA}}
\affil[3]{{\fontsize{12pt}{0.5em}\selectfont Health Analytics Network, Pennsylvania, USA}}

\newcommand{\ABSTRACT}{
The COVID-19 pandemic has had profound global consequences on health, economic, social, behavioral, and almost every major aspect of human life. Therefore, it is of great importance to model COVID-19 and other pandemics in terms of the broader social contexts in which they take place. 
We present the architecture of an artificial intelligence enhanced COVID-19 analysis (in short AICov) , which provides an integrative deep learning framework for COVID-19 forecasting with population covariates, some of which may serve as putative risk factors. 
We have integrated multiple different strategies into AICov, including the ability to use deep learning strategies based on LSTM and event modeling.
To demonstrate our approach, we have introduced a framework that integrates population covariates from multiple sources. Thus, AICov not only includes data on COVID-19 cases and deaths but, more importantly, the population's socioeconomic, health, and behavioral risk factors at a local level. The compiled data are fed into AICov, and thus we obtain improved prediction by the integration of the data to our model as compared to one that only uses case and death data. As we use deep learning our models adapt over time while learning the model from past data.
}
\date{}

@@

During outbreaks of epidemics, initially data on cases and deaths could be scarce, and the quality of data annotation, validation, and aggregation might be uncertain. For instance, changes in clinical data entry, such as the addition of a new category clinically diagnosed to the existing lab-confirmed category, may likely have been reflected in the reporting \cite{Petropoulos2020-hh}. Initial fears about COVID-19 among the general population with regards to the trajectory of the pandemic could also affect administrative reporting. Diverse media sources and differences in local and federal policies may add to the general uncertainty about disease progression. Nevertheless, a data-driven approach to forecasting can offer valuable insights into the disease dynamics, and thereby an ability to objectively plan for the near future.

Unlike earlier global outbreaks, during COVID-19, the current worldwide digital ecosystem allows for real-time data collection, which in turn is used by artificial intelligence (AI) and deep learning systems to understand healthcare trends, model risk associations, and predict outcomes \cite{Ting2020-lw}. 
In addition to traditional public health surveillance strategies available to most countries, a variety of static and dynamic data types may be integrated to model the scale and dynamics of this pandemic \cite{Pyne2015-ao}. Several organizations, including the Johns Hopkins University’s Center for Systems Science and Engineering, the New York Times, the Atlantic’s COVID-19 tracking project, among others have developed real-time tracking maps for following COVID-19 cases around the world using data from the U.S.~Centers for Disease Control and Prevention (CDC), WHO, and other international health agencies. Notably, the CDC receives forecasts from several modeling groups that use a wide variety of approaches ranging from SEIR (Susceptible-Exposed-Infectious-Recovered) to Agent-based to Bayesian models to understand the non-pharmaceutical policy interventions (or the lack thereof) to predict disease dynamics and its impact on human lives \cite{www-cdc-modeling-forecast}.

For precise and contextually relevant modeling of COVID-19 dynamics for a given population or community, we use data integration to combine community-specific health and behavioral risk factors, demographic and socioeconomic variables that are available at the local levels along with the cases and deaths data for COVID-19 while leveraging the recently developed and increasingly popular deep learning strategies to forecast COVID-19 dynamics using our model. Towards this, we develop a parallel computing platform called AICov standing for  AI-driven Platform for COVID-19 to implement the different components of our framework and obtain the needed resources through multi-cloud interfaces in a robust manner.

In this study, while we incorporate information from community and county-specific covariates to inform our forecasting model for each metropolitan area, we want to avoid the so-called ecological fallacy in drawing inferences about individual disease outcomes based on large area-level aggregated risk factors. Our objective is also to underscore the nuanced roles played by pre-existing socioeconomic and other prevailing conditions of a given community that can act as determinants of its health outcomes and possible disparities,  especially under such sudden stresses to its current systems as those felt during any pandemic.
It is important to note that the underlying models are learned directly from the data instead of being projected by a theoretical epidemiological model that relies on simulation populations. This allows quick adaptation and self-adaptation in regards to changing behaviors and regional differences.

@@

@@

@@

\begin{requirement}{\bf Expandable API.} Assuming that new models and other analysis algorithms will be developed over time, AICov must allow the integration of such APIs. \Solution To address this requirement, we are developed an abstraction API for data, analysis, and metadata parameter adjustments. External services could be integrated with the help of REST (REpresentational State Transfer) services. The REST services are automatically generated from function specifications. 
\end{requirement}

@@

Among the different sources of data on daily cases and deaths data collected and aggregated at different spatial levels, especially for the U.S., many are either freely available repositories such as GitHub or accessible via specifically developed online resources such as the New York Times~\cite{github-nytimes} and the Johns Hopkins Coronavirus Resource Center~\cite{www-jh-covid19}. 
For this study, we generated two data sets based on the latter resource upon our checking for the locations and times at which cumulative counts of COVID-19 cases and deaths on any given day were no less than those on the previous day. This led us to compile daily coronavirus cases and deaths data for 345 U.S. counties from 32 states and the District of Columbia and matched by five-digit FIPS code or county name to dynamic and static variables from additional data sources.

The static data in this repository was collected from the U.S. Census Bureau and the Centers for Disease Control and Prevention (CDC). The different fields cover the domains of behavioral and health risk factors, hospital capacity, and socioeconomic and demographic conditions. Data on health outcomes, prevention measures, and unhealthy behaviors were extracted from the CDC's 500 Cities Local Data for Better Health program \cite{www-cdc-chronic-data} in the form of crude or age-adjusted prevalence of conditions such as respiratory disease, obesity or smoking. These values were aggregated from the census tract to the county level using the first five digits of the eleven-digit tract FIPS code. Because a single county may consist of multiple individual cities, we include the list of all city labels within each aggregate group to represent a greater metropolitan area. 110 of such metropolitan areas that had more than 500 reported cases of COVID-19 by April 15, 2020, were selected for this study.

Notably, a greater metropolitan area such as New York City could be spread across more than one county, which can lead to the complexity of aggregating counts and other variables. Time series data for the 5 individual counties (boroughs) in New York City were not available in the Johns Hopkins CSSE (Center for Systems Science and Engineering) data set. Rather, the total number of cases and deaths for the entire city of New York are reported and assigned the FIPS code of New York County (Manhattan). To ensure consistent geography across the static and dynamic data, we compute the population-weighted sum or median of each covariate over the assigned counties.

@@

\begin{table}[!hptb]
\caption{Covariate Risk factor abbreviations as used in this study.}
\label{tab:risk-factors}
\bigskip
\resizebox{0.90\textwidth}{!}{
\centering
\begin{tabular}{ll}
\toprule
Risk Factor & Description \\
\midrule
 NONE & No risk factor used\\
 ARTHRITIS 	  & Percent reported with arthritis \\
 BINGE 		  & Percent reported with binge drinking \\
 BPHIGH 	  & Percent reported with  high blood pressure\\
 BPMED 		  & Percent reported taking blood pressure medication\\
 CANCER 	  & Percent reported as cancer patients\\
 CASTHMA 	  & Percent reported with current asthma\\
 CHD 		  & Percent reported with coronary heart disease\\
 CHECKUP 	  & Percent reported with health checkup\\
 CHOLSCREEN   & Percent reported with cholesterol screening\\
 COPD 		  & Percent reported with chronic obstructive pulmonary disease\\
 CSMOKING 	  & Percent reported as currently smoking\\
 DIABETES 	  & Percent reported with diabetes\\
 ESTBEDS 	  & Number of estimated beds in hospitals\\
 HIGHCHOL 	  & Percent reported with high cholesterol\\
 INSURANCE 	  & Percent reported with insurance\\
 KIDNEY 	  & Percent reported with kidney disease\\
 LPA 		  & Percent reported with no leisure-time physical activity \\
 MHLTH 		  & Percent reported with not good mental health\\
 NBEDS 		  & Number of beds in hospitals \\
 NBEDS/1000   & Number of beds per 1000 people\\
 NHOSP 		  & Number of hospitals\\
 OBESITY 	  & Percent reported with  population that is obese\\
 PHLTH        & Percent reported with physical health issues\\
 PVI 		  & Pandemic vulnerability index \\
 PERCENTBLACKS 	  & Percent of blacks in the population\\
 PERCENTHISPANICS & Percent of hispanics in the population\\
 STROKE 	  & Percent reported with stroke\\
 BLACK	  & Percent of Blacks in the population\\
 POVERTY & Percent of poverty in the population\\
 SENIOR  & Percent of seniors in the population\\
 SVI\_MINORITY 	  & Percent of social vulnerability index in the minority population\\
 SVI\_OVERALL 	  & Percent of social vulnerability index in the overall population\\
\bottomrule
\end{tabular}
}
\end{table}

@@
